---
title: "MESA_longitudinal_GP_lab_figures_manuscript&supp"
output: html_document
---


```{r}
library(ggplot2)
library(reshape2)
library(GGally)
library(dplyr)
library("PerformanceAnalytics")
```

```{r}
data <- read.csv("~/0_PhD/1st_year/3_rawdata/MESA data/labdatacorrected/mesalab4-12r.csv", header=TRUE)
```

Figure 2: 
Box Plots: Total Proportions of Hits By Month
```{r}
##longitudinal boxlots from 4 to 12 months
ggplot(data,aes(x=age, 
                y=uwAllpctStr, 
                group=age))+
  geom_boxplot()+
  geom_point(alpha=0.3)+
  labs(x='Age (months)', 
       y='Proportion Correct')+
  ylim(0,1)+
  geom_smooth(color='black')+
  theme_bw()+
  theme(
    panel.border = element_blank(), 
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), 
    axis.line = element_line(colour = "black")
    )+
  scale_x_continuous(breaks= c(4,6,8,10,12))+
  theme(text = element_text(size = 18))+
  coord_cartesian(ylim = c(0, 1))+
  geom_smooth(method = "lm", 
              se=TRUE, 
              aes(group=1), 
              show.legend = T)+
  #geom_jitter(alpha=0.4)+
  stat_sum(alpha=0.4)
```


Figure 3: Scatterplots: Hit Proportions By Month By Cue Type (Gaze, Point, Gaze+Point)
```{r}
#reshape for cue type
melted = melt(data, id.vars = c("ID", "age"),
                measure.vars = c("wtGzpctStr", "wtPTpctStr", "wtGPpctStr"))
```


```{r}
##longitudinal cue type
labels <- list(
  'Gaze', "Point", "Gaze + Point")
cuetype_labeller <- function(variable,value){
  return(labels[value])
}
ggplot(melted, aes(x=age, 
                   y=value, 
                   group=variable)) +
  geom_point(aes(x=age, 
                 y=value))+
  coord_cartesian(ylim = c(0, 1))+
  theme_bw()+
  theme(panel.grid.major =element_blank(),
        panel.grid.minor = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))+
  ylab("Proportion Correct")+
  xlab("Age (Months)")+
  scale_x_continuous(breaks=seq(4, 12, 2))+
  theme(text = element_text(size = 17),
        axis.text.x =element_text(size=17))+
  theme(axis.ticks.x=element_blank())+
  facet_wrap(~variable, 
             labeller=cuetype_labeller)+
  geom_smooth(method = "lm", 
              se=TRUE,
              show.legend = T)+
  geom_jitter(alpha=0.4)+
  stat_sum(alpha=0.4)
```


Figure 4: Scatterplots of Hit Proportions By Month By Target Location (Front, Side, Back)
```{r}
#reshape for cue location
melted1 = melt(data, id.vars = c("ID", "age"),
                measure.vars = c("wtFrpctStr", "wtSidepctStr", "wtBackpctStr"))
```

```{r}
##longitudinal cue location
labels <- list(
  'Front', "Side", "Back")
cuetype_labeller <- function(variable,value){
  return(labels[value])
}
ggplot(melted1, aes(x=age, 
                    y=value, 
                    group=variable)) +
  geom_point(aes(x=age, 
                 y=value))+
  coord_cartesian(ylim = c(0, 1))+
  theme_bw()+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))+
  ylab("Proportion Correct")+
  xlab("Age (Months)")+
  scale_x_continuous(breaks=seq(4, 12, 2))+
  theme(text = element_text(size = 17),
        axis.text.x = element_text(size=17))+
  theme(axis.ticks.x=element_blank())+
  facet_wrap(~variable, 
             labeller=cuetype_labeller)+
  geom_smooth(method = "lm", 
              se=TRUE,  
              show.legend = T)+
  geom_jitter(alpha = 0.4) +
  stat_sum(alpha=0.4)
```

Figure 5: Boxplots Of Hit Proportions By Main and Comparison Groups At 9 and 12 Months
```{r}
data1 <- read.csv("~/0_PhD/1st_year/3_rawdata/MESA data/labdatacorrected/mainsample_cs3_9&12.csv", header=TRUE)
```

```{r}
bxp <- ggboxplot(
  data1, x = "age", y = "hit",
  color = "group", palette = "jco"
  )
bxp
```


Figure 6: Correlations Of Hit Proportions From 7 To 12 Months
```{r}
###correlation among main sample
melted2 = melt(data, id.vars = c("ID", "age"),
                measure.vars = c("wtAllpctStr"))
melted2 = melted2[, c(1,2,4)]
head(melted2)
```

```{r}
bymonth= reshape(melted2, idvar = "ID", timevar = "age", direction = "wide")
bymonth= bymonth[, -1]
colnames(bymonth) = c( 'Mon4', 'Mon5',"Mon6", 'Mon7', 'Mon8', 'Mon9', 'Mon10', 'Mon11', 'Mon12' )
```


```{r}
my_fn <- function(data, mapping, ...){
    p <- ggplot(data = bymonth, mapping = mapping) + 
      geom_point() + 
      geom_smooth(method=lm, fill="blue", color="blue", ...)
  p
}

cor_func <- function(data, mapping, method, ...){
  x <- eval_data_col(data, mapping$x)
  y <- eval_data_col(data, mapping$y)
  
  corr <- cor(x, y, method=method, use='complete.obs')
  
  if(stars){
  stars <- c("***", "**", "*", "")[findInterval(corr$p.value, c(0, 0.001, 0.01, 0.05, 1))]
  lbl <- paste0(round(est, ndp), stars)
}else{
  lbl <- round(est, ndp)
}

  
  ggally_text(
    label = as.character(round(corr, 2)), 
    mapping = aes(),
    xP = 0.25, yP = 0.25,
    color = 'black',
    ...
  )
} 

cor_fun <- function(data, mapping, method="pearson", use="pairwise", ndp=2, sz=9, stars=TRUE, ...){

# grab data
x <- eval_data_col(data, mapping$x)
y <- eval_data_col(data, mapping$y)

# calculate correlation: for significance stars
corr <- cor.test(x, y, method=method)
est <- corr$estimate
lb.size <- 4

# get significance stars
if(stars){
  stars <- c("***", "**", "*", "")[findInterval(corr$p.value, c(0, 0.001, 0.01, 0.05, 1))]
  lbl <- paste0(round(est, ndp), stars)
}else{
  lbl <- round(est, ndp)
}

# calculate correlation: for colored tiles
corr <- cor(x, y, method=method, use=use)

# calculate color based on correlation value
# corr = -1 => blue, 
# corr =  0 => white, 
# corr = +1 => red, 
colFn <- colorRampPalette(c("blue", "white", "red"), interpolate ='spline')
fill <- colFn(100)[findInterval(corr, seq(-1, 1, length=100))]

ggplot(data = data, mapping = mapping, ...) + 
  theme_void() +
  annotate("text",
           x=mean(x, na.rm=TRUE),
           y=mean(y, na.rm=TRUE),
           label=lbl,
           size=lb.size,
           ...) 
  
}
###plot correlations for month 7- 12
#g = ggpairs(bymonth, 
#            columns=4:9, 
#            lower = list(continuous = my_fn),
#            upper=list(continuous=cor_fun))+
#  theme(text= element_text(size = 15))

b = ggpairs(bymonth, 
            lower = list(continuous = my_fn),
            upper=list(continuous=cor_fun,
                       continuous = wrap("cor", size=20) ),
            columns=4:9)+
  theme(text = element_text(size=13))+
  scale_x_continuous(guide = guide_axis(check.overlap = TRUE))+
  scale_y_continuous(guide = guide_axis(check.overlap = TRUE))
b
```


----------------------------Supplemental figures---------------------------------

Figure S1: Boxplots of Liberal and Stringent Hits, By Infant Age and Target Location

```{r}
###correlation between liberal and strigent
strlib = read.csv("~/0_PhD/1st_year/3_rawdata/MESA data/labdatacorrected/mesa_lab_GP_strlib.csv", header=TRUE)
strlib = strlib[strlib$location != "all",]
strlib
```

```{r}
libdata <- read.csv("~/0_PhD/1st_year/3_rawdata/MESA data/labdatacorrected/mesalab4-12libr.csv", header=TRUE)
```


```{r}
bxp <- ggboxplot(
  strlib, x = "age", y = "hitpct", 
  color = "location", facet.by = "criterion"
  )
bxp
```




Figure S3: Correlations Between AF at 9-10 and 11-12 Months, And Language Development Indices
```{r}
###9-12month and language scores (unweighted)
data4 <- read.csv("~/0_PhD/1st_year/3_rawdata/MESA data/9-12lablanguage.csv", header=TRUE)
data4 <- na.omit(data4)
langwt <-data4[, c(7, 15, 16, 17, 18, 19, 20, 21)]
languw <-data4[, c(14, 15, 16, 17, 18, 19, 20, 21)]
langsupp <- data4[, c(5, 6,  15, 16, 17, 18, 19, 20, 21)]
head(langsupp)
```


```{r}
##correlation table for language scores (supplemental)
b <- ggpairs(langsupp, showStrips = 'TRUE',diag = list(continuous = "densityDiag"))

#b[1,1] = b[1,1] + scale_y_continuous(limits = c(0,10))
cor_func <- function(data, mapping, method, ...){
  x <- eval_data_col(data, mapping$x)
  y <- eval_data_col(data, mapping$y)
  
  corr <- cor(x, y, method=method, use='complete.obs')
  
  if(stars){
  stars <- c("***", "**", "*", "")[findInterval(corr$p.value, c(0, 0.001, 0.01, 0.05, 1))]
  lbl <- paste0(round(est, ndp), stars)
}else{
  lbl <- round(est, ndp)
}

  
  ggally_text(
    label = as.character(round(corr, 2)), 
    mapping = aes(),
    xP = 0.25, yP = 0.25,
    color = 'black',
    ...
  )
 
}

cor_fun <- function(data, mapping, method="pearson", use="pairwise", ndp=2, sz=5, stars=TRUE, ...){

# grab data
x <- eval_data_col(data, mapping$x)
y <- eval_data_col(data, mapping$y)

# calculate correlation: for significance stars
corr <- cor.test(x, y, method=method)
est <- corr$estimate
lb.size <- 4

# get significance stars
if(stars){
  stars <- c("***", "**", "*", "")[findInterval(corr$p.value, c(0, 0.001, 0.01, 0.05, 1))]
  lbl <- paste0(round(est, ndp), stars)
}else{
  lbl <- round(est, ndp)
}

# calculate correlation: for colored tiles
corr <- cor(x, y, method=method, use=use)

# calculate color based on correlation value
# corr = -1 => blue, 
# corr =  0 => white, 
# corr = +1 => red, 
colFn <- colorRampPalette(c("blue", "white", "red"), interpolate ='spline')
fill <- colFn(100)[findInterval(corr, seq(-1, 1, length=100))]

ggplot(data = data, mapping = mapping, ...) + 
  theme_void() +
  annotate("text",
           x=mean(x, na.rm=TRUE),
           y=mean(y, na.rm=TRUE),
           label=lbl,
           size=lb.size,
           ...) 
  
}

my_fn <- function(data, mapping, ...){
  p <- ggplot(data = data, mapping = mapping) + 
    geom_point() + 
    
    geom_smooth(method=lm, fill="blue", color="blue", ...)
  p
}


b = ggpairs(langsupp, 
        upper=list(continuous=cor_fun, 
                   continuous = wrap("cor", size=20)),
        lower = list(continuous = my_fn), 
        columnLabels = c("Ave 9-10", "Ave 11-12", "EC12", "RC12", "Comp12","EC18", "RC18", "Comp18", "MCDI22Ave" ))+
  theme(text = element_text(size=12)) +
  scale_x_continuous(guide = guide_axis(check.overlap = TRUE))+
  scale_y_continuous(guide = guide_axis(check.overlap = TRUE))

b
```



Figure S4: correlation among main sample from 4 to 12 months
```{r}

###plot correlations for month 4- 12
#g = ggpairs(bymonth, lower = list(continuous = my_fn),upper=list(continuous=cor_fun))+theme(text= #element_text(size = 15))
#g

b = ggpairs(bymonth, 
            lower = list(continuous = my_fn),
            upper=list(continuous=cor_fun,
                       continuous = wrap("cor", size=20) ))+
  theme(text = element_text(size=15))+
  scale_x_continuous(guide = guide_axis(check.overlap = TRUE))+
  scale_y_continuous(guide = guide_axis(check.overlap = TRUE))
b
```

Figure S5: Linear Slopes From 4 To 9 To 12 Months For The Main And Comparison Groups

```{r}
datanew1 <- read.csv("~/0_PhD/1st_year/3_rawdata/MESA data/labdatacorrected/mainsample_cs3_growth.csv", header=TRUE)
```

```{r}
ggplot() +
  geom_smooth(aes(x = age, y = hit), data = datanewcsnine1, 
              method = "lm", se = FALSE, color = "#D16103") + 
  geom_smooth(aes(x = age, y = hit), data = datanewmstotal, 
              method = "lm", se = FALSE, color = "#4E84C4") + 
  geom_point(aes(x = age, y = hit), data = datanewcsnine1, color = "#D16103") + 
  geom_point(aes(x = age, y = hit), data = datanewmstotal, color = "#4E84C4")+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"),text = element_text(size=20))

```
