---
title: "MESA_longitudinal_GP_lab_stats"
output: html_document
---


```{r}
library(ggplot2)
library(reshape2)
library(GGally)
library(dplyr)
library("PerformanceAnalytics")
library(tidyverse)
library(ggpubr)
library(rstatix)
library(afex)
library(lme4)
library(nlme)
```

```{r}
###longitudinal GP main sample stringent data
data <- read.csv("~/0_PhD/1st_year/3_rawdata/MESA data/labdatacorrected/mesalab4-12r.csv", header=TRUE)
```

```{r}
#orientation data
oridata <- read.csv("~/0_PhD/1st_year/3_rawdata/MESA data/mesa_lab_4-6orientationtrial.csv", header=TRUE)

```

```{r}
#liberal data
libdata <- read.csv("~/0_PhD/1st_year/3_rawdata/MESA data/labdatacorrected/mesalab4-12libr.csv", header=TRUE)
```

```{r}
head(oridata)
```


```{r}
six = as.numeric(dfRefined$sixori)
sd(six[complete.cases(six)])
```

```{r}
### visualize distribution of 4, 5, 6 ori data
#hist(oridata$fourori)
fourori=oridata$fourori[!is.na(oridata$fourori)]
fiveori=oridata$fiveori[!is.na(oridata$fiveori)]
hist(fourori)
median(fiveori)
median(fourori)
### for four month, we have a roughtly normal distribution, with similar mean and median, take 4 as cutoff
### for month five, we have a left skewed distribution, with median at 5 and mean 4.9, take 5 as cutoff
sum(fiveori>=5)
length(fiveori)
sum(fourori>=4)
length(fourori)
```

```{r}
###pair ori data with hit data
prdfour = oridata[which(!is.na(oridata$fourori) & !is.na(oridata$fourhit)),]
prdfive = oridata[which(!is.na(oridata$fiveori) & !is.na(oridata$fivehit)),]
fourori1 = prdfour$fourori
fiveori1 = prdfive$fiveori
hist(fiveori1)
length(fourori1) ##21 for four month
sum(fourori1>=4)
length(fiveori1) ##33 for five month
sum(fiveori1>=5)
### hit rates for four month, divide groups based on oridata
group1four = prdfour[prdfour$fourori>=4,]$fourhit
group2four = prdfour[prdfour$fourori<4,]$fourhit
###hit rates for five month
group1five = prdfive[prdfive$fiveori>=5,]$fivehit
group2five = prdfive[prdfive$fiveori<5,]$fivehit
```

```{r}
###test
wilcox.test(group1four, group2four, alternative = "two.sided")
wilcox.test(group1five, group2five, alternative = "two.sided")
```


```{r}
#difference between 4&5 orientation data
wilcox.test(oridata$fourori, oridata$fiveori, alternative = "two.sided")
t.test(oridata$fourori, oridata$fiveori, alternative = "two.sided")
```

```{r}
#4&5
dfRefined=oridata[which(!is.na(oridata$fourori) & !is.na(oridata$fiveori)),]
test=wilcox.test(dfRefined$fourori,dfRefined$fiveori,paired=TRUE)
zscore = qnorm(test$p.value/2)
zscore
```

```{r}
##5&6
dfRefined=oridata[which(!is.na(oridata$sixori) & !is.na(oridata$fiveori)),]
test=wilcox.test(as.numeric(dfRefined$sixori),dfRefined$fiveori,paired=TRUE)
zscore = qnorm(test$p.value/2)
zscore
test
```


```{r}
#difference between 4&6 orientation data

dfRefined=oridata[which(!is.na(oridata$fourori) & !is.na(oridata$sixori)),]
test=wilcox.test(dfRefined$fourori,as.numeric(dfRefined$sixori),paired=TRUE)
zscore = qnorm(test$p.value/2)
zscore
test
```

```{r}
#predicting 4 month/correlation between ori data and hit rate
new2=oridata[which(!is.na(oridata$fourori) & !is.na(oridata$fourhit)),]
res2 <- cor.test(new$fourori, new$fourhit, 
                    method = "pearson")
res2
```

```{r}
#predicting 5 month
new=oridata[which(!is.na(oridata$fiveori) & !is.na(oridata$fivehit)),]
res <- cor.test(new$fiveori, new$fivehit, 
                    method = "pearson")
res
```

```{r}
#predicting 6 month
new1=oridata[which(!is.na(oridata$sixori) & !is.na(oridata$sixhit)),]
res1 <- cor.test(new1$sixori, new$sixhit, 
                    method = "pearson")
res1
```

```{r}
### ----------orientation data over, continue with testing phase-----###
```



```{r}
### check for main sample difference between 4&5
datatemp = data[data$age==4 | data$age==5,]
fourmonth = datatemp[datatemp$age==4,]
fourmonth = fourmonth$wtAllpctStr
fivemonth = datatemp[datatemp$age==5,]
fivemonth = fivemonth$wtAllpctStr
test = wilcox.test(fourmonth,fivemonth, alternative = "two.sided")
zscore = qnorm(test$p.value/2)
zscore
```


```{r}
### check for main sample difference between 4&6
datatemp = data[data$age==4 | data$age==6,]
fourmonth = datatemp[datatemp$age==4,]
fourmonth = fourmonth$wtAllpctStr
sixmonth = datatemp[datatemp$age==6,]
sixmonth = sixmonth$wtAllpctStr
test = wilcox.test(fourmonth,sixmonth, alternative = "two.sided")
test
zscore = qnorm(test$p.value/2)
zscore
```

```{r}
### check for main sample difference between 5&6
datatemp = data[data$age==5 | data$age==6,]

sixmonth = datatemp[datatemp$age==6,]
sixmonth = sixmonth$wtAllpctStr
test = wilcox.test(fivemonth,sixmonth, alternative = "two.sided")
test
zscore = qnorm(test$p.value/2)
zscore
```

```{r}
t.test(fivemonth, fourmonth)
```




```{r}
#check for sex differences
###wilcoxon ranked sum test
test = wilcox.test(data$wtAllpctStr~data$sex, alternative = "two.sided")
test
zscore = qnorm(test$p.value/2)
zscore
```


```{r}
hist(data$uwAllpctStr)
```


Effects of sex on AF performance (nonsignificant)
```{r}
#check for interaction between sex and cue type (weighted)
#two way anova
# Gather the columns unwtgaze, unwtpoint and unwtGP into long format.
# Convert sex and age into factor variables
data1 <- data %>%
  gather(key = "cuetype", value = "hitrate", wtGzpctStr, wtPTpctStr, wtGPpctStr) %>%
  convert_as_factor(sex,cuetype )
# Inspect some random rows of the data by groups
#set.seed(123)
data1
res.aov2 <- aov(hitrate ~ sex + cuetype+sex:cuetype, data = data1)
summary(res.aov2)
```



Stringent vs Liberal criterion
```{r}
###linear mixed effect model on stringent data from 4 to 12 months; weighted
#scaled.data <- scale(data) stringent data
lmm <- lmer(scale(wtAllpctStr) ~ scale(age) + (1 | ID), data = data)
summary(lmm)
```

```{r}
###linear mixed effect model on liberal data from 4 to 12 months; unweighted
#scaled.data <- scale(data) liberal data
lmm1 <- lmer(scale(AllpctLib) ~ scale(age) + (1 | ID), data = libdata)
summary(lmm1)
```



Effects of cue type
```{r}
#comparison between GP and pointing
test=wilcox.test(data$wtGPpctStr, data$wtPTpctStr, paired = TRUE, alternative = "two.sided")
test
zscore=qnorm(test$p.value/2)
zscore
```

```{r}
#comparison between GP and gaze
test=wilcox.test(data$wtGPpctStr, data$wtGzpctStr, paired = TRUE, alternative = "two.sided")
test
zscore=qnorm(test$p.value/2)
zscore
```

```{r}
#comparison between pointing and gaze
wilcox.test(data$wtPTpctStr, data$wtGzpctStr, paired = TRUE, alternative = "two.sided")
```

Effects of cue location
```{r}
#comparison between front and back
test=wilcox.test(data$wtFrpctStr, data$wtBackpctStr, paired = TRUE, alternative = "two.sided")
zscore=qnorm(test$p.value/2)
zscore
test
```

```{r}
#comparison between front and side
test=wilcox.test(data$wtFrpctStr, data$wtSidepctStr, paired = TRUE, alternative = "two.sided")
zscore=qnorm(test$p.value/2)
zscore
test
```

```{r}
#comparison between side and back
wilcox.test(data$wtSidepctStr, data$wtBackpctStr, paired = TRUE, alternative = "two.sided")
```

Cue type developmental trajectory
```{r}
#overall developmental trajectory for gp
lm(uwGPpctStr ~ age, data=data)

```

```{r}
#overall developmental trajectory for gaze
lm(uwGzpctStr ~ age, data=data)
```

```{r}
#individual developmental trajectory for GP
devGP = coef(lmList(wtGPpctStr~age | ID, data=data))$age
#individual developmental trajectory for POINT
devPT = coef(lmList(wtPTpctStr ~age| ID, data=data))$age
#individual developmental trajectory for GAZE
devGz = coef(lmList(wtGzpctStr~age | ID, data=data))$age
```

```{r}
devcue = data.frame(devGP, devPT, devGz)
devcue <- devcue %>%
  gather(key = "cuetype", value = "hitrate", devGP, devPT, devGz) %>%
  convert_as_factor(cuetype )
```

```{r}
res.aov <- aov(hitrate ~ cuetype, data = devcue)
summary(res.aov)
```

```{r}
###developmental trajetory comparison between gaze and pointing 
t.test(devGz, devPT)
```


```{r}
###developmental trajetory comparison between GP and pointing 
t.test(devGP, devPT)
```

```{r}
###developmental trajetory comparison between GP and gaze 
t.test(devGP, devGz)
```

```{r}
hist(devGP)
```



Developmental trajectory of different target locations AF
```{r}
#individual developmental trajectory for fr
devfr = coef(lmList(wtFrpctStr~age | ID, data=data))$age
#individual developmental trajectory for side
devsd = coef(lmList(wtSidepctStr ~age| ID, data=data))$age
#individual developmental trajectory for bk
devbk = coef(lmList(wtBackpctStr~age | ID, data=data))$age
```


```{r}
###developmental trajetory comparison between front and side
t.test(devfr, devsd)


###developmental trajetory comparison between front and back 
t.test(devfr, devbk)


###developmental trajetory comparison between side and back 
t.test(devsd, devbk)
```


```{r}
devloc = data.frame(devfr, devsd, devbk)
devloc <- devloc %>%
  gather(key = "cueloc", value = "hitrate", devfr, devsd, devbk) %>%
  convert_as_factor(cueloc )
devloc
```

```{r}
res.aov <- aov(hitrate ~ cueloc, data = devloc)
summary(res.aov)
```

```{r}
hist(devGz)
```


Comparison between strigent and liberal criterion
```{r}
###correlation between strigent and liberal hits by month
corstrlib = data <- read.csv("~/0_PhD/1st_year/3_rawdata/MESA data/labdatacorrected/mesa_lab_GP_strlib.csv", header=TRUE)

```


```{r}
library(dplyr)
corstrlib %>%
  group_by(age) %>%
  summarize(COR=cor(AllpctLib,Allpctstr))
```


Correlation between trial number and AF (data retention)
```{r}
###correlation between trial number and hit (weighted stringent overall)
cortrialhit = data <- read.csv("~/0_PhD/1st_year/3_rawdata/MESA data/labdatacorrected/mesalab_trial_3moave.csv", header=TRUE)
firstthree = cortrialhit[cortrialhit$age==4,]
cor.test(firstthree$avetrial, firstthree$avehit)
```

```{r}
secondthree = cortrialhit[cortrialhit$age==7,]
cor.test(secondthree$avetrial, secondthree$avehit)
```

```{r}
lastthree = cortrialhit[cortrialhit$age==10,]
cor.test(lastthree$avetrial, lastthree$avehit)
```


```{r}
###correlation between session number and hit
session = data <- read.csv("~/0_PhD/1st_year/3_rawdata/MESA data/labdatacorrected/mesalab_sessiondata.csv", header=TRUE)

library(dplyr)
session %>%
  group_by(age) %>%
  summarize(COR=cor(Allpctstr,missing))
```

```{r}
cor.test(session$Allpctstr, session$missing)
```

```{r}
###correlation between liberal and strigent
strlib = read.csv("~/0_PhD/1st_year/3_rawdata/MESA data/labdatacorrected/mesa_lab_GP_strlib.csv", header=TRUE)
strlib = strlib[strlib$location != "all",]
strlib
```

```{r}
library(dplyr)

#calculate mean hit increase grouped by age, criterion, and location
strlib %>%
  group_by(age, criterion, location) %>%
  summarize(mean_hit = mean(hitpct))
```

```{r}
strlib$age <- as.factor(strlib$age)
strlib$location <- as.factor(strlib$location)
strlib$criterion <- as.factor(strlib$criterion)
```


```{r}
#perform three-way ANOVA
model <- aov(hitpct ~ age * criterion * location, data=strlib)

#view summary of three-way ANOVA
summary(model)

```

```{r}
library(dplyr)

#find mean jumping increase by program
strlib %>%
  group_by(criterion) %>%
  summarize(mean_hit = mean(hitpct))


```

```{r}
bxp <- ggboxplot(
  strlib, x = "age", y = "hitpct", 
  color = "location", facet.by = "criterion"
  )
bxp
```

practice effects: main sample and comparison sample
```{r}
datanew <- read.csv("~/0_PhD/1st_year/3_rawdata/MESA data/labdatacorrected/mainsample_cs3_9&12.csv", header=TRUE)
```

```{r}
###month 9 data; comparison between cs and ms
datanewcs9 = datanewcs[datanewcs$age==9,]$hit
datanewms9 = datanewms[datanewms$age==9,]$hit
test = wilcox.test(datanewms9, datanewcs9)
zscore=qnorm(test$p.value/2)
zscore
test
```

```{r}
###month 12 data; comparison between cs and ms
datanewcs12 = datanewcs[datanewcs$age==12,]$hit
datanewms12 = datanewms[datanewms$age==12,]$hit
test = wilcox.test(datanewms12, datanewcs12)
zscore=qnorm(test$p.value/2)
zscore
test
```

```{r}
#perform ANOVA
model <- aov(hit ~ age * factor(group), data=datanew)
summary(model)
```


```{r}
datanewcs = datanew[datanew$group=='cs',]
#cs3 growth curve 9-12
lmm2 <- lmer(scale(hit) ~ scale(age) + (1 |id ), data = datanewcs)
summary(lmm2)
```

```{r}
datanewms = datanew[datanew$group=='ms',]
#main sample growth curve 9-12
lmm3 <- lmer(scale(hit) ~ scale(age) + (1 |id ), data = datanewms)
summary(lmm3)
```

```{r}
datanew1 <- read.csv("~/0_PhD/1st_year/3_rawdata/MESA data/labdatacorrected/mainsample_cs3_growth.csv", header=TRUE)
```

```{r}
#main sample 4-12
datanewmstotal = datanew1[datanew1$group=='ms',]
#cs3 growth curve 9-12
lmm3 <- lmer(scale(hit) ~ scale(age) + (1 |id ), data = datanewmstotal)
summary(lmm3)
```

```{r}
#main sample 4-9

databewmsnine = datanewmstotal[datanewmstotal$age==4 |datanewmstotal$age==9 ,]
#cs3 growth curve 9-12
lmm2 <- lmer(scale(hit) ~ scale(age) + (1 |id ), data = datanewmsnine)
summary(lmm2)
```

```{r}
#cs sample 4-9
#databewcsnine= datanew1[datanew1$group=='cs',]
datanewcsnine = datanew1[(datanew1$age==4 &datanew1$group=='ms')|(datanew1$age==9 &datanew1$group=='cs') ,]
datanewcsnine
lmm4 <- lmer(scale(hit) ~ scale(age), data = datanewcsnine)
summary(lmm4)
```

```{r}
#cs sample 4-9-12
#databewcsnine= datanew1[datanew1$group=='cs',]
datanewcsnine1 = datanew1[(datanew1$age==4 &datanew1$group=='ms')|(datanew1$age==9 &datanew1$group=='cs')|(datanew1$age==12 &datanew1$group=='cs') ,]
#cs3 growth curve 9-12
lmm5 <- lmer(scale(hit) ~ scale(age) + (1 |id ), data = datanewcsnine1)
summary(lmm5)
```

```{r}
ggplot() +
  geom_smooth(aes(x = age, y = hit), data = datanewcsnine1, 
              method = "lm", se = FALSE, color = "#D16103") + 
  geom_smooth(aes(x = age, y = hit), data = datanewmstotal, 
              method = "lm", se = FALSE, color = "#4E84C4") + 
  geom_point(aes(x = age, y = hit), data = datanewcsnine1, color = "#D16103") + 
  geom_point(aes(x = age, y = hit), data = datanewmstotal, color = "#4E84C4")+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"),text = element_text(size=20))

```
